---
output: github_document
bibliography: bib.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Reproduce the Results of @gerard2020pairwise

## Introduction

This repository contains the code and instructions needed to reproduce all of
the results from @gerard2020pairwise.

If you find a bug, please create an
[issue](https://github.com/dcgerard/reproduce_pairwise_ld/issues).

## Instructions

You'll need to have up-to-date versions of R and python3 to run these simulatoins.
    
1. Install the appropriate R packages
    ```{r, eval = FALSE}
    install.packages(c("tidyverse",
                       "ggthemes",
                       "latex2exp",
                       "Rcpp",
                       "RcppArmadillo",
                       "foreach",
                       "doParallel",
                       "corrplot",
                       "devtools"))
    devtools::install_github("dcgerard/updog")
    devtools::install_github("dcgerard/ldsep")
    ```
    
2. Install ngsLD in the working directory of the simulation files: <https://github.com/fgvieira/ngsLD>
    
    ``` bash
    git clone https://github.com/fgvieira/ngsLD.git
	  cd ngsLD
	  make
    ```

3. Install msprime: <https://msprime.readthedocs.io/en/stable/installation.html>
  
    I did this via pip:
    
    ``` bash
    pip3 install msprime
    ```
    
    Make sure that you have added `~/.local/bin` to the `PATH` by adding the
    following to your `.bashrc` file:
    
    ``` bash
    export PATH=$PATH:~/.local/bin
    ```
    
3. Download the following files from Uitdewilligen et. al. (2013), 
   place them in the "data" folder, and extract them.
    
    - https://doi.org/10.1371/journal.pone.0062355.s004
    - https://doi.org/10.1371/journal.pone.0062355.s007
    
    You can do this with the following bash code
    
    ``` bash
    sudo apt-get install p7zip p7zip-full p7zip-rar
    wget --directory-prefix=data -nc https://doi.org/10.1371/journal.pone.0062355.s004
    mv ./data/journal.pone.0062355.s004 ./data/journal.pone.0062355.s004.gz
    wget --directory-prefix=data -nc https://doi.org/10.1371/journal.pone.0062355.s007
    mv ./data/journal.pone.0062355.s007 ./data/journal.pone.0062355.s007.gz
    7z e ./data/journal.pone.0062355.s004.gz -o./data/
    7z e ./data/journal.pone.0062355.s007.gz -o./data/
    ```
    
    As long as you have 7z installed, `make` will attempt to download and extract the files for you.

4. Run `make`. This will run all of the simulations.

    ``` bash
    make
    ```
  
  You may choose to run only part of hte simulations `make mle` or `make ngsLD`.

5. Get coffee/sweets. Running `make sims` should take a few hours. You should get 
   some coffee! Here is a list of some of my favorite places:

    - Washington, DC
        - [Colony Club](https://www.yelp.com/biz/colony-club-washington)
        - [Three Fifty](https://www.yelp.com/biz/three-fifty-bakery-and-coffee-bar-washington) (for the baked goods, coffee is meh)
    - Chicago
        - [Sawada Coffee](https://www.yelp.com/biz/sawada-coffee-chicago)
        - [Plein Air Cafe](https://www.yelp.com/biz/plein-air-cafe-and-eatery-chicago-2)
    - Seattle
        - [Bauhaus Ballard](https://www.yelp.com/biz/bauhaus-ballard-seattle)
        - [Cafe Solstice](https://www.yelp.com/biz/cafe-solstice-seattle)
    - Columbus
        - [Yeah, Me Too](https://www.yelp.com/biz/yeah-me-too-columbus)
        - [Stauf's Coffee Roasters](https://www.yelp.com/biz/staufs-coffee-roasters-columbus-2)


Note that I've also only tried this on Ubuntu.

## References
    
- Uitdewilligen, Jan GAML, Anne-Marie A. Wolters, Bjorn B. D’hoop, Theo JA Borm, Richard GF Visser, and Herman J. van Eck. "A next-generation sequencing method for genotyping-by-sequencing of highly heterozygous autotetraploid potato." _PloS one_ 8, no. 5 (2013).
