---
output: github_document
bibliography: bib.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Reproduce the Results of @gerard2020pairwise

## Introduction

This repository contains the code and instructions needed to reproduce all of
the results from @gerard2020pairwise.

If you find a bug, please create an
[issue](https://github.com/dcgerard/reproduce_pairwise_ld/issues).

## Instructions
    
1. Install the appropriate R packages
    ```{r, eval = FALSE}
    install.packages(c("updog",
                       "tidyverse", 
                       "Rcpp",
                       "RcppArmadillo",
                       "roptim",
                       "foreach",
                       "doParallel",
                       "devtools"))
    devtools::install_github("dcgerard/ldsep")
    devtools::install_github("tpbilton/GUSbase")
    devtools::install_github("AgResearch/GUS-LD")
    ```

6. Get coffee. Running `make sims` should take a few hours. You should get 
   some coffee! Here is a list of some of my favorite places:

    - Washington, DC
        - [Colony Club](https://www.yelp.com/biz/colony-club-washington)
        - [Grace Street Coffee](https://www.yelp.com/biz/grace-street-coffee-georgetown)
    - Chicago
        - [Sawada Coffee](https://www.yelp.com/biz/sawada-coffee-chicago)
        - [Plein Air Cafe](https://www.yelp.com/biz/plein-air-cafe-and-eatery-chicago-2)
    - Seattle
        - [Bauhaus Ballard](https://www.yelp.com/biz/bauhaus-ballard-seattle)
        - [Cafe Solstice](https://www.yelp.com/biz/cafe-solstice-seattle)
    - Columbus
        - [Yeah, Me Too](https://www.yelp.com/biz/yeah-me-too-columbus)
        - [Stauf's Coffee Roasters](https://www.yelp.com/biz/staufs-coffee-roasters-columbus-2)


Note that I've also only tried this on Ubuntu.

## References
    

