---
output: github_document
bibliography: bib.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Reproduce the Results of @gerard2021pairwise

[![DOI](https://img.shields.io/badge/doi-10.6084/m9.figshare.12765803-blue.svg?style=flat&labelColor=whitesmoke&logo=data%3Aimage%2Fpng%3Bbase64%2CiVBORw0KGgoAAAANSUhEUgAAAB8AAAAfCAYAAAAfrhY5AAAJsklEQVR42qWXd1DTaRrHf%2BiB2Hdt5zhrAUKz4IKEYu9IGiGFFJJQ0gkJCAKiWFDWBRdFhCQUF3UVdeVcRQEBxUI3yY9iEnQHb3bdW1fPubnyz%2F11M7lvEHfOQee2ZOYzPyDv%2B3yf9%2Fk95YX4fx%2BltfUt08GcFEuPR4U9hDDZ%2FVngIlhb%2FSiI6InkTgLzgDcgfvtnovhH4BzoVlrbwr55QnhCtBW4QHXnFrZbPBaQoBh4%2FSYH2EnpBEtqcDMVzB93wA%2F8AFwa23XFGcc8CkT3mxz%2BfXWtq9T9IQlLIXYEuHojudb%2BCM7Hgdq8ydi%2FAHiBXyY%2BLjwFlAEnS6Jnar%2FvnQVhvdzasad0eKvWZKe8hvDB2ofLZ%2FZEcWsh%2BhyIuyO5Bxs2iZIE4nRv7NWAb0EO8AC%2FWPxjYAWuOEX2MSXZVgPxzmRL3xKz3ScGpx6p6QnOx4mDIFqO0w6Q4fEhO5IzwxlSwyD2FYHzwAW%2BAZ4fEsf74gCumykwNHskLM7taQxLYjjIyy8MUtraGhTWdkfhkFJqtvuVl%2F9l2ZquDfEyrH8B0W06nnpH3JtIyRGpH1iJ6SfxDIHjRXHJmdQjLpfHeN54gnfFx4W9QRnovx%2FN20aXZeTD2J84hn3%2BqoF2Tqr14VqTPUCIcP%2B5%2Fly4qC%2BUL3sYxSvNj1NwsVYPsWdMUfomsdkYm3Tj0nbV0N1wRKwFe1MgKACDIBdMAhPE%2FwicwNWxll8Ag40w%2BFfhibJkGHmutjYeQ8gVlaN%2BjO51nDysa9TwNUFMqaGbKdRJZFfOJSp6mkRKsv0rRIpEVWjAvyFkxNOEpwvcAVPfEe%2Bl8ojeNTx3nXLBcWRrYGxSRjDEk0VlpxYrbe1ZmaQ5xuT0u3r%2B2qe5j0J5uytiZPGsRL2Jm32AldpxPUNJ3jmmsN4x62z1cXrbedXBQf2yvIFCeZrtyicZZG2U2nrrBJzYorI2EXLrvTfCSB43s41PKEvbZDEfQby6L4JTj%2FfIwam%2B4%2BwucBu%2BDgNK05Nle1rSt9HvR%2FKPC4U6LTfvUIaip1mjIa8fPzykii23h2eanT57zQ7fsyYH5QjywwlooAUcAdOh5QumgTHx6aAO7%2FL52eaQNEShrxfhL6albEDmfhGflrsT4tps8gTHNOJbeDeBlt0WJWDHSgxs6cW6lQqyg1FpD5ZVDfhn1HYFF1y4Eiaqa18pQf3zzYMBhcanlBjYfgWNayAf%2FASOgklu8bmgD7hADrk4cRlOL7NSOewEcbqSmaivT33QuFdHXj5sdvjlN5yMDrAECmdgDWG2L8P%2BAKLs9ZLZ7dJda%2BB4Xl84t7QvnKfvpXJv9obz2KgK8dXyqISyV0sXGZ0U47hOA%2FAiigbEMECJxC9aoKp86re5O5prxOlHkcksutSQJzxZRlPZmrOKhsQBF5zEZKybUC0vVjG8PqOnhOq46qyDTDnj5gZBriWCk4DvXrudQnXQmnXblebhAC2cCB6zIbM4PYgGl0elPSgIf3iFEA21aLdHYLHUQuVkpgi02SxFdrG862Y8ymYGMvXDzUmiX8DS5vKZyZlGmsSgQqfLub5RyLNS4zfDiZc9Edzh%2FtCE%2BX8j9k%2FqWB071rcZyMImne1SLkL4GRw4UPHMV3jjwEYpPG5uW5fAEot0aTSJnsGAwHJi2nvF1Y5OIqWziVCQd5NT7t6Q8guOSpgS%2Fa1dSRn8JGGaCD3BPXDyQRG4Bqhu8XrgAp0yy8DMSvvyVXDgJcJTcr1wQ2BvFKf65jqhvmxXUuDpGBlRvV36XvGjQzLi8KAKT2lYOnmxQPGorURSV0NhyTIuIyqOmKTMhQ%2BieEsgOgpc4KBbfDM4B3SIgFljvfHF6cef7qpyLBXAiQcXvg5l3Iunp%2FWv4dH6qFziO%2BL9PbrimQ9RY6MQphEfGUpOmma7KkGzuS8sPUFnCtIYcKCaI9EXo4HlQLgGrBjbiK5EqMj2AKWt9QWcIFMtnVvQVDQV9lXJJqdPVtUQpbh6gCI2Ov1nvZts7yYdsnvRgxiWFOtNJcOMVLn1vgptVi6qrNiFOfEjHCDB3J%2BHDLqUB77YgQGwX%2Fb1eYna3hGKdlqJKIyiE4nSbV8VFgxmxR4b5mVkkeUhMgs5YTi4ja2XZ009xJRHdkfwMi%2BfocaancuO7h%2FMlcLOa0V%2FSw6Dq47CumRQAKhgbOP8t%2BMTjuxjJGhXCY6XpmDDFqWlVYbQ1aDJ5Cptdw4oLbf3Ck%2BdWkVP0LpH7s9XLPXI%2FQX8ws%2Bj2In63IcRvOOo%2BTTjiN%2BlssfRsanW%2B3REVKoavBOAPTXABW4AL7e4NygHdpAKBscmlDh9Jysp4wxbnUNna3L3xBvyE1jyrGIkUHaqQMuxhHElV6oj1picvgL1QEuS5PyZTEaivqh5vUCKJqOuIgPFGESns8kyFk7%2FDxyima3cYxi%2FYOQCj%2F%2B9Ms2Ll%2Bhn4FmKnl7JkGXQGDKDAz9rUGL1TIlBpuJr9Be2JjK6qPzyDg495UxXYF7JY1qKimw9jWjF0iV6DRIqE%2B%2FeWG0J2ofmZTk0mLYVd4GLiFCOoKR0Cg727tWq981InYynvCuKW43aXgEjofVbxIqrm0VL76zlH3gQzWP3R3Bv9oXxclrlO7VVtgBRpSP4hMFWJ8BrUSBCJXC07l40X4jWuvtc42ofNCxtlX2JH6bdeojXgTh5TxOBKEyY5wvBE%2BACh8BtOPNPkApjoxi5h%2B%2FFMQQNpWvZaMH7MKFu5Ax8HoCQdmGkJrtnOiLHwD3uS5y8%2F2xTSDrE%2F4PT1yqtt6vGe8ldMBVMEPd6KwqiYECHDlfbvzphcWP%2BJiZuL5swoWQYlS%2Br7Yu5mNUiGD2retxBi9fl6RDGn4Ti9B1oyYy%2BMP5G87D%2FCpRlvdnuy0PY6RC8BzTA40NXqckQ9TaOUDywkYsudxJzPgyDoAWn%2BB6nEFbaVxxC6UXjJiuDkW9TWq7uRBOJocky9iMfUhGpv%2FdQuVVIuGjYqACbXf8aa%2BPeYNIHZsM7l4s5gAQuUAzRUoT51hnH3EWofXf2vkD5HJJ33vwE%2FaEWp36GHr6GpMaH4AAPuqM5eabH%2FhfG9zcCz4nN6cPinuAw6IHwtvyB%2FdO1toZciBaPh25U0ducR2PI3Zl7mokyLWKkSnEDOg1x5fCsJE9EKhH7HwFNhWMGMS7%2BqxyYsbHHRUDUH4I%2FAheQY7wujJNnFUH4KdCju83riuQeHU9WEqNzjsJFuF%2FdTDAZ%2FK7%2F1WaAU%2BAWymT59pVMT4g2AxcwNa0XEBDdBDpAPvgDIH73R25teeuAF5ime2Ul0OUIiG4GpSAEJeYW9wDTf43wfwHgHLKJoPznkwAAAABJRU5ErkJggg%3D%3D)](https://doi.org/10.6084/m9.figshare.12765803)

## Introduction

This repository contains the code and instructions needed to reproduce all of
the results from @gerard2021pairwise.

Some of the plots from the simulations mentioned in the manuscript can be found [here](https://github.com/dcgerard/ld_simulations/tree/master/output/mle_plots) and [here](https://github.com/dcgerard/ld_simulations/tree/master/output/comp/comp_plots).

If you find a bug, please create an
[issue](https://github.com/dcgerard/ld_simulations/issues).

## Instructions

You will need to have R and GNU Make installed to run these simulations. There are other dependencies for [ngsLD](https://github.com/fgvieira/ngsLD), [PedigreeSim](https://github.com/PBR/pedigreeSim), and some of the R packages I used. Check their documentation if you are having trouble installing them.

Note that I have only tried these out on `r version$version.string` running `r sessionInfo()$running`.
    
1. Install the appropriate R packages
    ```{r, eval = FALSE}
    install.packages("BiocManager")
    BiocManager::install(c("tidyverse",
                           "ggthemes",
                           "gridExtra",
                           "latex2exp",
                           "Rcpp",
                           "RcppArmadillo",
                           "foreach",
                           "doParallel",
                           "corrplot",
                           "devtools",
                           "vcfR",
                           "VariantAnnotation"))
    devtools::install_github("dcgerard/updog")
    devtools::install_github("dcgerard/ldsep")
    ```
    
```{r, echo = FALSE, eval = TRUE, message = FALSE, warning=FALSE, results='hide'}
pkgvec <- c("BiocManager", 
            "tidyverse",
            "ggthemes",
            "gridExtra",
            "latex2exp",
            "Rcpp",
            "RcppArmadillo",
            "foreach",
            "doParallel",
            "corrplot",
            "devtools",
            "vcfR",
            "VariantAnnotation",
            "updog",
            "ldsep")
lapply(pkgvec, require, character.only = TRUE)
```
    
    
2. Install ngsLD [@fox2019ngsld] in the working directory of the simulation files: <https://github.com/fgvieira/ngsLD>
    
    ``` bash
    git clone https://github.com/fgvieira/ngsLD.git
    cd ngsLD
    make
    ```
    
3. Download the following files from @uitdewilligen2013next, 
   place them in the "data" folder, and extract them.
    
    - <https://doi.org/10.1371/journal.pone.0062355.s004>
    - <https://doi.org/10.1371/journal.pone.0062355.s007>
    - <https://doi.org/10.1371/journal.pone.0062355.s009>
    - <https://doi.org/10.1371/journal.pone.0062355.s010>
    
    You can do this with the following bash code from the top directory of the ld_simulations repo.
    
    ``` bash
    wget --directory-prefix=data -nc https://doi.org/10.1371/journal.pone.0062355.s004
    mv ./data/journal.pone.0062355.s004 ./data/journal.pone.0062355.s004.gz
    wget --directory-prefix=data -nc https://doi.org/10.1371/journal.pone.0062355.s007
    mv ./data/journal.pone.0062355.s007 ./data/journal.pone.0062355.s007.gz 
    wget --directory-prefix=data -nc https://doi.org/10.1371/journal.pone.0062355.s009
    mv ./data/journal.pone.0062355.s009 ./data/journal.pone.0062355.s009.xls 
    wget --directory-prefix=data -nc https://doi.org/10.1371/journal.pone.0062355.s010
    mv ./data/journal.pone.0062355.s010 ./data/journal.pone.0062355.s010.xls 
    7z e ./data/journal.pone.0062355.s004.gz -o./data/
    7z e ./data/journal.pone.0062355.s007.gz -o./data/
    ```
    
    As long as you have 7z installed, `make` will attempt to download and extract the files for you.
    
4. Download the data from @mcallister2016single <https://doi.org/10.5061/dryad.05qs7> and unzip it to the ./data/gerardii directory. I unzipped using 7z:
    
    ``` bash
    7z e doi_10.5061_dryad.05qs7__v1.zip 
    ```
    
5. Your data directory should now look like this:

    ```{r, echo = FALSE, comment=""}
    writeLines(list.files("./data", recursive = TRUE, full.names = TRUE))
    ```

6. Run `make`. This will run all of the analyses.

    ``` bash
    make
    ```
  
    You may choose to run only part of the analyses via:
    
    - `make mle`: Simulations under HWE.
    - `make comp`: Simulations under violations from HWE.
    - `make norm`: Visualizing the proportional bivariate normal distribution.
    - `make ddiff`: Comparing, under HWE, D' and the Δ' that conditions on the marginal genotype distributions.
    - `make ngsLD`: Verifying that [`ldsep`](https://cran.r-project.org/package=ldsep) and [`ngsLD`](https://github.com/fgvieira/ngsLD) [@fox2019ngsld] provide the same results in diploids.
    - `make uit`: Real-data analysis using the data from @uitdewilligen2013next.
    - `make mca`: Real-data analysis using the data from @mcallister2016single.
    - `make ped`: Simulations under *interpretable* violations from HWE, where data were simulated using [PedigreeSim](https://github.com/PBR/pedigreeSim) [@voorrips2012simulation].

7. Get coffee/sweets. Running `make sims` should take a few hours. You should get some coffee! Here is a list of some of my favorite places:

    - Washington, DC
        - [Doubles](https://www.yelp.com/biz/doubles-washington)
        - [Little Red Fox](https://www.yelp.com/biz/little-red-fox-washington)
        - [Three Fifty](https://www.yelp.com/biz/three-fifty-bakery-and-coffee-bar-washington) (for the baked goods, coffee is meh)
    - Chicago
        - [Sawada Coffee](https://www.yelp.com/biz/sawada-coffee-chicago)
        - [Plein Air Cafe](https://www.yelp.com/biz/plein-air-cafe-and-eatery-chicago-2)
    - Seattle
        - [Bauhaus Ballard](https://www.yelp.com/biz/bauhaus-ballard-seattle)
        - [Cafe Solstice](https://www.yelp.com/biz/cafe-solstice-seattle)
    - Columbus
        - [Yeah, Me Too](https://www.yelp.com/biz/yeah-me-too-columbus)
        - [Stauf's Coffee Roasters](https://www.yelp.com/biz/staufs-coffee-roasters-columbus-2)

# Session Information

```{r, echo = FALSE}
sessionInfo()
```

## References
